Keystoneness, centrality, and the structural controllability of ecological networks

### 2018-12-24 paper/reviewers-response.Rmd

# Associate Editor's comments:

> "I think this ms now makes much clearer the potential insights to be gained by applying control theory to interaction networks. In particular, the authors have done a nice job of explaining how their findings reinforce existing ideas in the literature on the stability of ecological networks while also acknowledging limitations. The reviewers identify a few remaining issues that should be straightforward to address. I also noticed a few issues"

Thank you for the positive feedback. We have addressed all of the specific issues raised and highlight these in red text in our resubmitted manuscript.

> L9: I think here and throughout much of the text, species should be possessive when followed by a noun ("species role" should be "species' role"). The possessive is currently used in some instances but is inconsistently applied throughout the ms.

We have checked through the manuscript and corrected this where relevant.

> L25: ". every invasive species in our dataset was part of it ." Here it was unclear what "it" refers to (the set of critical species), and the reference to a singular dataset when you have 20 networks in total was a bit confusing to me.

We have made this clearer as follows: "and included every invasive species in our dataset" (\lr{R2-our-dataset})

> L23-26: in these lines, you refer to critical species, but I think the distinction between having high control capacity versus being critical needs to be defined (briefly) earlier (it is stated in L18-19 that you will quantify both control capacity and identify  critical species, but it is not explained how they are distinct).

We now identify the distinction between high control capacity and being critical on lines \lr{R2-define-control-capacity} "It also allows us to quantify a species control capacity-its relative importance-and identify the set of species that, collectively, are critical-have maximum control capacity-in this context."

> L27: unclear what "its" refers  to

We now explicitly state "species' degree" \lr{R2-species-degree}

> L111, 114: "are comprised of" should be "comprised "

Thank you, we have corrected this.

> L234, 235: awkward to refer to species as "they" and "their "

We have changed this to "it" and "its"

> L296: "More details in Supporting Information ."  is a sentence fragment.

We have corrected this on line \lr{R2-sentence-fragment}

> L339, 432: "at" should be "in"

Thank you, this has been corrected.

> L341: "allow" should be "allowed "

Thank you, this has been corrected.

> L343: I am not sure about citing Table 2 for the first time in the Discussion. Might it be more useful to cite this in the Introduction ?

We now introduce table 2 on line \lr{R2-ref-glossary}.

>L355: missing period

Thank you for noticing this, we have inserted the missing period.

@@

# Reviewer comments #1

> "I'd like to commend the authors to this revised version of the manuscript on "Keystoneness, centrality, and the structural controllability of ecological networks". I thoroughly enjoyed the re-written manuscript, which is now considerably more accessible and clearly targets the community of network ecologists, while not ignoring the importance of the applied relevance of the findings. Overall, I don't have any major comments. I guess the language is in parts on the casual side, yet without compromising the preciseness or clarity of the text. I noticed a few typos or style errors that the authors may want to address"

We are pleased that the reviewer "thoroughly enjoyed the re-written manuscript", and appreciate the identified issues. We have addressed all issues in the text and identify these in red.

> A few references require editing to journal style (e.g. remove initials for many citations)

>L314. open bracket missing .

Thank you for pointing this out, we have inserted the missing bracket.

> L358. change 'instead' to 'Instead'. L381. remove second comma

We have made both of these changes.

> L388. remove repeated 'other'. I suggest that the authors re-write this sentence to improve clarity .

We have removed the second 'other'. We have rewritten the sentence on line \lr{R2-structural-controllability} to improve clarity. It now reads as follows: "First, structural controllability emphasizes that the effect a species has on the abundance of other species is not independent of the effects of those other species in their community".

> Reference list needs updating, e.g. Ne'eman et al. is not 2009 but 2010.

> Fig. S6 header of plot b). correct spelling of 'critical'

> Congratulations on this thought-provoking methodological synthesis.

@@

# Reviewer comments #2

> I think the concepts presented in this manuscript are a useful addition to the literature on management of ecological networks.  I agree with the authors that it is important that the discussion surrounding what constitutes a "keystone" species should go beyond centrality metrics.

> I found the results and concepts to be generally well presented, particularly graphically, and the authors present a thorough sensitivity analysis to various assumptions.  The methods are mostly clear but there is no description of how the "likelihood of being a superior node " is calculated.

> I would suggest including clarification on the limitations of using control capacity to summarise species' importance, and examples to help the reader build an intuition regarding the control capacity metric.  As pointed out, networks that give rise to a species having a high control capacity can be quite distinct to networks that give rise to a species having a high value of particular centrality metrics.  Providing an example in the Discussion of, for instance, those species that have a control capacity of 1 but a low degree would be very useful.

> Jargon occasionally obfuscates the concepts being introduced and there are places where I thought the language used was vague (esp. if terms being used have particular meanings depending on the field you're coming from).  I would suggest clarifying, or rewording, the following :

> * "desired control trajectory " - does this mean "final state"?

We have attempted to clarify this sentence as follows: Finding out exactly how difficult it is to control a network depends strongly on the particularities of the desired control trajectory (i.e. path to the desired final state)  as well as the dynamical relationship between nodes. \lr{R2-desired-control-trajectory}

> * "dynamic persistence"  (p17,l364)

> * "dynamic stability " (p17, l370)

> * "ecological objective" (p9) - simply 'objective' ?

We have changed this to simply 'objective', as suggested (\lr{R2-ecol-objective}).

> * "central in their community " (p18, l 405)

> * "inherently different " (p18, l 406). The sentence spanning lines 407-410 is quite confusing  - please clarify.

We rephrased this line as follows: "We found that it is not that invasive plants have some different mechanism for influencing the community compared to their native counterparts" \lr{R2-inherrently-different}.

> * "marginally affected by the topology of their interactions"  (p19, l 430)

We have rephrased this line as: "By accounting for network dynamics (even if in a simple way), structural stability incorporates more ecological realism, especially in the extreme scenario in which the structure of interactions within the community only marginally affects the community's state. Both native species and invasive plants tended to attain a high control capacity if they were important to network persistence, were abundant, and depended little on other species." \lr{R2-marginally-affected}.

> * "dominate the state [of their community ]" (p19, l436 & l441)

>* "species' strength " (Discussion)

>Minor typos:
* Figure S11 caption: "likelyhood" -> likelihood 
* References : several references use the term "plantPollinator", or "..."
* Figure S1b is referred to (supp. p4 ) but doesn't exist in the figure.  
* Page 18 Line 388: "other other" 
* "driver node density" (supp p12) -> "size of the minimum driver node set"?

### 2018-12-24 paper/supp-info.Rmd
further updates to revision, and some minor corrections to the supplementary info

The networks studied had species richness ranging between `r min(net_prop$n_s_t)` and `r max(net_prop$n_s_t)` when considering only the largest component in each network). 
As shown by the network asymmetry *AS* [@bluthgen_specialization_2007], the networks had a low ratio of plants to pollinators overall. 
Furthermore, the networks had relatively low levels of nestedness [when measured using the quantitative version of the NODF index; @almeida-neto_straightforward_2011]. Details for each network can be found in the Table \@ref(tab:table-network-properties).

```{r table-network-properties, results = "asis", sanitize = FALSE}
net_prop %>%
    dplyr::select(-net_name, -count, -net) %>%
    dplyr::filter(study != "ballantyne") %>%
    dplyr::arrange(study, site, inv) %>% 
    dplyr::mutate(invader = dplyr::case_when(
                     invader == "car" ~ "\\textit{Carpobrotus affine}",
                     invader == "op" ~ "\\textit{Opuntia stricta}",
                     invader == "imp" ~ "\\textit{Impatients grandulifera}",
                     TRUE ~ "—"
                 ),
                 invader = dplyr::if_else(inv, invader, "—")) %>%
    dplyr::mutate(site = rep(1:n_distinct(site), each = 2),
                 inv = dplyr::if_else(inv, "yes", "no"),
                 study = dplyr::case_when(
                     study == "bartomeus" ~ "Cap de Creus, Spain",
                     TRUE ~ "Bristol, United Kingdom"
                 )) %>% 
  dplyr::select(site, invader, n_sp, n_pla, n_pol, connectance, web.asymmetry, weighted.NODF, study) %>%
  dplyr::rename_(
    # "$R$" = "n_s_t",
                    "$n_p$" = "n_pla",
                    "$n_a$" = "n_pol",
                    "location" = "study",
                    "$n_s$" = "n_sp",
                    "$c$"= "connectance", 
                    "NODF" = "weighted.NODF",
                    # "$\\bar{a}$" = "interaction.strength.asymmetry",
                    "$AS$" = "web.asymmetry") %>% 
  knitr::kable("latex", 
               escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c("", "\\addlinespace"),
               caption = "Properties of the analysed plant-pollinator communities. Here show the number of species ($n_s$), the number of plants ($n_p$), the number of pollinators ($n_a$), the network connectance ($c$), the network asymmetry ($AS$), and the network nestedness (NODF index). All properties correspond to the network's largest component. British networks were assembled by Lopezaraiza-Mikel et al. (2007), Spanish were networks assembled by Bartomeus et al. (2008).") %>%
  kableExtra::kable_styling(font_size = 8)
```

@@

The algorithm proposed by @zhang_input_2016 enable us to find all minimum driver node sets without calculating all possible maximum matchings. 
The algorithm is based on the construction of the input graph of the directed network. 
Components in the input graph reveal the correlations between nodes from a structural control perspective (Figure \@ref(fig:fig-input-graph)). 
The input graph is constructed based on the *control adjacency* of nodes in a maximum matching. 
Two nodes $x_i$ and $x_j$ are said to be control adjacent if there is a node $x_k$ connecting $x_i$ and $x_j$ with an unmatched link $x_k-x_i$ and a matched link $x_k-x_j$. 
For example $a_1$ is control adjacent to $a_2$ in the top panel of Figure \@ref(fig:fig-input-graph). 
If a node $x_i$ is part of a minimum driver node set then every node in the same component of $x_i$ in the input graph is a possible driver node.

@@

In order to calculate a node's control capacity $\phi$, we harness the fact that nodes can be classified into three groups according to their role in the input graph. 
First,  if $x_i$ belongs to the minimum driver node set and has no adjacent nodes, then it cannot be replaced and will be part of every possible control configuration. 
Second, if $x_i$ does not belong to the minimum driver node set and is not adjacent to any possible input node, then it does not part of the minimum driver node set under any control configuration. 
Finally, if a node $x_i$ belongs to the minimum driver node set and is adjacent to another node $x_j$ in the input graph, then $x_j$ must belong to another minimum driver node to which $x_i$ does not belong, in other words, $x_i$ and $x_j$ are substitutable.

The first and the second group have a control capacity of $\phi = 1$ and $\phi = 0$ respectively. 
Meanwhile, the third group has intermediate values of control capacity which can be calculated by computing the number of possible substitutions of a node by its control adjacent nodes. 
As a rule of thumb, the largest the number of possible substitutions an input node has, the smaller its control capacity.

@@

At a network scale the driver-node density $n_D$ was consistent among the three weighting schemes (`r round(bal_nd$eff, 2)` for deposition, `r round(bal_nd$vis, 2)` for the visitation, and `r round(bal_nd$imp, 2)` for the pollinator-importance network, respectively).
The choice of weighting can also have an impact on at the species level.
Therefore we calculated $\phi$ and $\sigma$ and calculated it's correlation among all three weighting schemes.
Although visitation and efficiency (pollen deposition) produce moderately different results, we found a very strong agreement between the order produced by visitation and importance which is arguably a more accurate metric of interspecific effects (Table \@ref(tab:tab-assumption-visitation)).

@@

To do so we removed a portion of the visits, for each network, and calculated how three control metrics of the subsampled network compare to those of the full network. 
Specifically, we calculated the difference between the relative size of the minimum driver node set $n_D$ of the subsampled and the full network, the Spearman correlation between the control capacities $\phi$ and the probability of being a superior node $\sigma$ that were obtained for the empirical network and subsamples of it. 
Finally, we also compared the control capacity of critical species ($\phi=1$) across the different subsamples. 
We removed up to 20% of the interactions in 1% increments. 
We repeated the procedure for each network to obtain a total sample size of $n=400$.

Overall, the results of the sensitivity analysis indicate that our approach is likely to still be useful in the absence of complete sampling. 
Specifically, as the proportion of sampled interaction decreased, the variability of $n_D$ increased but was overall similar to that obtained when using all available interactions (Figure \@ref(fig:fig-assumption-sampling-nd)). 
Similarly, the correlation between $\phi$ and $\sigma$ of empirical and subsampled networks was high even for extreme levels of undersampling (Figure \@ref(fig:fig-assumption-sampling-cor)). 
Importantly, critical species in the empirical networks were also critical species in the subsampled networks in a large majority of cases (Figure \@ref(fig:fig-assumption-sampling-crit)).

@@

c_mod_table %>% 
  dplyr::mutate(weight = round(weight, 2)) %>%
  knitr::kable("latex", 
               escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c(""),
               caption = "Model selection table based on the Akaike Information Criterion for small samples for candidate models of size of the minimum driver node set ($n_D$) in the empirical networks. There was no evidence for any of the evaluated predictors: the invasion status (inv), the network connectance (c), the network nestedness (NODF) the number of species ($n_S$) the network asymmetry ($AS_s$) or the interaction strength asymmetry ($AS_s$). Only candidate models with a weight $>0.01$ are shown.") %>%
  kableExtra::kable_styling(font_size = 8)
```

>>>>

The networks studied had species richness ranging between `r min(net_prop$n_s_t)` and `r max(net_prop$n_s_t)` (when considering only the largest component in each network). 
As shown by the network asymmetry *AS* [@bluthgen_specialization_2007], the networks had a low ratio of plants to pollinators overall. 
Furthermore, the networks had relatively low levels of nestedness [when measured using the quantitative version of the NODF index; @almeida-neto_straightforward_2011]. Details for each network can be found in the Table \@ref(tab:table-network-properties).

```{r table-network-properties, results = "asis", sanitize = FALSE}
net_prop %>%
    dplyr::select(-net_name, -count, -net) %>%
    dplyr::filter(study != "ballantyne") %>%
    dplyr::arrange(study, site, inv) %>% 
    dplyr::mutate(invader = dplyr::case_when(
                     invader == "car" ~ "\\textit{Carpobrotus affine}",
                     invader == "op" ~ "\\textit{Opuntia stricta}",
                     invader == "imp" ~ "\\textit{Impatients grandulifera}",
                     TRUE ~ "—"
                 ),
                 invader = dplyr::if_else(inv, invader, "—")) %>%
    dplyr::mutate(site = rep(1:n_distinct(site), each = 2),
                 inv = dplyr::if_else(inv, "yes", "no"),
                 study = dplyr::case_when(
                     study == "bartomeus" ~ "Cap de Creus, Spain",
                     TRUE ~ "Bristol, United Kingdom"
                 )) %>% 
  dplyr::select(site, invader, n_sp, n_pla, n_pol, connectance, web.asymmetry, weighted.NODF, study) %>%
  dplyr::rename_(
    # "$R$" = "n_s_t",
                    "$n_p$" = "n_pla",
                    "$n_a$" = "n_pol",
                    "location" = "study",
                    "$n_s$" = "n_sp",
                    "$c$"= "connectance", 
                    "NODF" = "weighted.NODF",
                    # "$\\bar{a}$" = "interaction.strength.asymmetry",
                    "$AS$" = "web.asymmetry") %>% 
  knitr::kable("latex", 
               escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c("", "\\addlinespace"),
               caption = "Properties of the analysed plant-pollinator communities. Here show the number of species ($n_s$), the number of plants ($n_p$), the number of pollinators ($n_a$), the network connectance ($C$), the network asymmetry ($AS$), and the network nestedness (NODF index). All properties correspond to the network's largest component. British networks were assembled by Lopezaraiza-Mikel et al. (2007), Spanish were networks assembled by Bartomeus et al. (2008).") %>%
  kableExtra::kable_styling(font_size = 8)
```

@@

The algorithm proposed by @zhang_input_2016 enables us to find all minimum driver node sets without calculating all possible maximum matchings. 
The algorithm is based on the construction of the input graph of the directed network. 
Components in the input graph reveal the correlations between nodes from a structural control perspective (Figure \@ref(fig:fig-input-graph)). 
The input graph is constructed based on the *control adjacency* of nodes in a maximum matching. 
Two nodes $x_i$ and $x_j$ are said to be control adjacent if there is a node $x_k$ connecting $x_i$ and $x_j$ with an unmatched link $x_k-x_i$ and a matched link $x_k-x_j$. 
For example $a_1$ is control adjacent to $a_2$ in the top panel of Figure \@ref(fig:fig-input-graph). 
If a node $x_i$ is part of a minimum driver node set then every node in the same component of $x_i$ in the input graph is a possible driver node.

@@

In order to calculate a node's control capacity $\phi$, we harness the fact that nodes can be classified into three groups according to their role in the input graph. 
First,  if $x_i$ belongs to the minimum driver node set and has no adjacent nodes, then it cannot be replaced and will be part of every possible control configuration. 
Second, if $x_i$ does not belong to the minimum driver node set and is not adjacent to any possible input node, then it does not become part of the minimum driver node set under any control configuration. 
Finally, if a node $x_i$ belongs to the minimum driver node set and is adjacent to another node $x_j$ in the input graph, then $x_j$ must belong to another minimum driver node to which $x_i$ does not belong, in other words, $x_i$ and $x_j$ are substitutable.

The first and the second group have a control capacity of $\phi = 1$ and $\phi = 0$ respectively. 
Meanwhile, the third group has intermediate values of control capacity which can be calculated by computing the number of possible substitutions of a node by its control adjacent nodes. 
As a rule of thumb, the larger the number of possible substitutions an input node has, the smaller its control capacity.

@@

At a network scale the driver-node density $n_D$ was consistent among the three weighting schemes (`r round(bal_nd$eff, 2)` for deposition, `r round(bal_nd$vis, 2)` for the visitation, and `r round(bal_nd$imp, 2)` for the pollinator-importance network, respectively).
The choice of weighting can also have an impact on at the species level.
Therefore we calculated $\phi$ and $\sigma$ and calculated its correlation among all three weighting schemes.
Although visitation and efficiency (pollen deposition) produce moderately different results, we found a very strong agreement between the order produced by visitation and importance which is arguably a more accurate metric of interspecific effects (Table \@ref(tab:tab-assumption-visitation)).

@@

To do so, we removed a portion of the visits, for each network, and calculated how three control metrics of the subsampled network compare to those of the full network. 
Specifically, we calculated the difference between the relative size of the minimum driver node set $n_D$ of the subsampled and the full network, the Spearman correlation between the control capacities $\phi$ and the probability of being a superior node $\sigma$ that were obtained for the empirical network and subsamples of it. 
Finally, we also compared the control capacity of critical species ($\phi=1$) across the different subsamples. 
We removed up to 20% of the interactions in 1% increments. 
We repeated the procedure for each network to obtain a total sample size of $n=400$.

Overall, the results of the sensitivity analysis indicate that our approach is likely to still be useful in the absence of complete sampling. 
Specifically, as the proportion of sampled interactions decreased, the variability of $n_D$ increased but was overall similar to that obtained when using all available interactions (Figure \@ref(fig:fig-assumption-sampling-nd)). 
Similarly, the correlation between $\phi$ and $\sigma$ of empirical and subsampled networks was high even for extreme levels of undersampling (Figure \@ref(fig:fig-assumption-sampling-cor)). 
Importantly, critical species in the empirical networks were also critical species in the subsampled networks in a large majority of cases (Figure \@ref(fig:fig-assumption-sampling-crit)).

@@

c_mod_table %>% 
  dplyr::mutate(weight = round(weight, 2)) %>%
  knitr::kable("latex", 
               escape = FALSE, 
               booktabs = TRUE, 
               digits = 2, 
               linesep = c(""),
               caption = "Model selection table based on the Akaike Information Criterion for small samples for candidate models of size of the minimum driver node set ($n_D$) in the empirical networks. There was no evidence for any of the evaluated predictors: the invasion status (inv), the network connectance ($C$), the network nestedness (NODF) the number of species ($n_S$) the network asymmetry ($AS_s$) or the interaction strength asymmetry ($AS_s$). Only candidate models with a weight $>0.01$ are shown.") %>%
  kableExtra::kable_styling(font_size = 8)
```
